---
title: "Shiny"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Shiny}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE,
  message=FALSE
)
```

### Load the package 

```{r setup, include=FALSE}

devtools::load_all()
#library(RShinyHelpers)

```

```{r setupshow, include=TRUE}
library(RShinyHelpers)

```

Call the Example Shiny app to demonstrate the shiny functionality using `demoApp()`

## Here we demonstrate the **value/info** boxes 

```{r finance, include=TRUE, echo= TRUE }

finance_vbox(toplot = -(1:10+runif(n=10,min=-1,max=1)),label = 'market index',textid = '£1M',col = 'black')


```



```{r vbox 1}

vbox(label = 'label',textid = 'value')

```



```{r 4vbox}

four_value(c('BCT','ETH','KLM','DOGE'),c('$4.00','£5.34','€7.47','£9.68'))

```


## SliderInput colour

We style the sliderInput with `bootstrap_slider_colour(slider_index = 0,colour = 'mediumseagreen')`


## DT 

To improve the aesthetics of the wildly popular and infinitely functional `DT::Datatables()`, based on the _jquery_ library of the same name, we implement some improved primary colouring.  
Note these functions now support manipulation of both bootstrap 3, using shiny's default, and bs4 datatables, if using bslib.

We set the hover colour _pink_ and selected colour _red_, with the supplementary text surrounding the bulk content, to a fun pinky-red _#ff003b_.

``` {r echo=FALSE, include=TRUE}

#ui <- fluidPage(theme=bslib::bs_theme(),
   
    ##call inside the ui declaration 

    DT_selected_row_colour(colour = 'red')
    DT_hover_row_colour(colour = 'pink')
    DT_peripheral_colour(colour = '#ff003b')
# ,..

```

```{r DT, include=TRUE }

DT::datatable(mtcars[1:5],selection = TRUE)


```

